---
title: "Course Project"
author: "Andrea Di Iura"
date: "30 novembre 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Course Project

### Load data

We load data from the directory `data`.

```{r}
training_name = "data/pml-training.csv"
testing_name = "data/pml-testing.csv"

training = read.csv(training_name, na.strings=c("", "NA", "#DIV/0!"))
testing = read.csv(testing_name, na.strings=c("", "NA", "#DIV/0!"))
```

Using the commands `head`, `str` and `summary` we investigate the training dataset, the output is quite long thus we do not show it here. The datasets have dimensions 

```{r}
dim(training)
dim(testing)
```

We can reduce the number of columns in the training dataset because the first column contains the indeces. We also remove the columns with all NA values. To do this we create a mask variable called `mask_training`.

```{r}
training$X = NULL
testing$X = NULL
```

```{r}
mask_training = colSums(is.na(training))<nrow(training)
training = training[,mask_training]
testing = testing[,mask_training]
```

We reduce the dimensions
```{r}
dim(training)
dim(testing)
```

## Exploratory analysis

The first step is to do some exploratory analysis to explore possible correlations among the variables.



